---
title: "Bariatric_surgery"
author: "Yihan Shi"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load library, cache = TRUE}
library(survival)
library(survminer)
library(ranger)
library(gt)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggfortify)
library(devtools)
library(OIsurv)
library(lubridate)
library(ciTools) # for accelerated failure time model
```

```{r load data, cache = TRUE}
data <- load("recur.RData")
data
```

## Data dictionary
In *x_outcome* dataset, 
Claim-level information - only claims which correspond to an alcohol-related outcome
alc: alcohol outcomes
FROM_DT: date of the claim (alcohol outcome), outcome for each individual patient (PAT_ID) can be tracked through date

*If days are consecutive, count as 1 occurrence*

*b_enroll* dataset contains patient level characteristics
# index = fist relevavnt surgery date
```{r bariatric surgery}
b_outcom # 1361 unique rows
b_expsum
b_enroll # 48997 rows
```

*c_expsum* dataset contains patient level characteristics for the year prior to the indexed surgery date
sums of how many times patients had a claim for various conditions (may adjust for baseline)

```{r control surgery }
c_outcom
c_expsum
c_enroll
```

## Cleaning
```{r cleaning b}
# Make a data set with unique combinations of patient ID & FROM_DT for outcom
b_outcom_unique <- b_outcom %>% 
  distinct(PAT_ID, FROM_DT, .keep_all = TRUE)
 
# this dataset only contains patient ID and occurrences 
b_outcom_longer <- b_outcom_unique %>% 
  group_by(PAT_ID) %>% 
  mutate(row_n = row_number()) %>% 
  pivot_wider(PAT_ID, names_from = row_n, values_from = FROM_DT, 
              names_glue = "Occurrence{row_n}") 
  
# check NA values in a column = 0
sum(is.na(b_outcom_unique$FROM_DT))

# Join enroll, outcom, and expsum
b_outcome_enroll <- b_enroll %>% 
  left_join(b_outcom_unique, c("pat_id" = "PAT_ID"))

b_all <- b_outcome_enroll %>% 
  left_join(b_expsum, c("pat_id" = "PAT_ID")) %>% 
  mutate(der_age = 2021 - der_yob,
         alc.x = replace_na(alc.x, 0),
         alcabuse.x = replace_na(alcabuse.x, 0),
         alcdep.x = replace_na(alcdep.x, 0),
         alcgastr.x = replace_na(alcgastr.x, 0),
         ald.x = replace_na(ald.x, 0),
         alcpois.x = replace_na(alcpois.x, 0),
         alcsuic.x = replace_na(alcsuic.x, 0),
         alcblood.x = replace_na(alcblood.x, 0),
         # create time (days) until event happen
         time = case_when(
           !is.na(FROM_DT) ~ as.numeric(difftime(FROM_DT, index.x, units = c("days"))), 
           is.na(FROM_DT) ~ as.numeric(difftime(clm_last, index.x,units = c("days")))), 
         surgery_type = "b")

# sum(b_enroll$pat_id %in% b_all$pat_id)

# dim(b_outcom_unique)
# dim(b_outcom)
```


```{r cleaning c}
c_outcom_unique <- c_outcom %>% 
  distinct(PAT_ID, FROM_DT, .keep_all = TRUE)
  
c_outcom_longer <- c_outcom_unique %>% 
  group_by(PAT_ID) %>% 
  mutate(row_n = row_number()) %>% 
  pivot_wider(PAT_ID, names_from = row_n, values_from = FROM_DT, 
              names_glue = "Occurrence{row_n}")

# Join enroll, outcom, and expsum
c_outcome_enroll <- c_enroll %>% 
  left_join(c_outcom_unique, c("pat_id" = "PAT_ID"))

c_all <- c_outcome_enroll %>% 
  left_join(c_expsum, c("pat_id" = "PAT_ID")) %>% 
  mutate(der_age = 2021 - der_yob,
         alc.x = replace_na(alc.x, 0),
         alcabuse.x = replace_na(alcabuse.x, 0),
         alcdep.x = replace_na(alcdep.x, 0),
         alcgastr.x = replace_na(alcgastr.x, 0),
         ald.x = replace_na(ald.x, 0),
         alcpois.x = replace_na(alcpois.x, 0),
         alcsuic.x = replace_na(alcsuic.x, 0),
         alcblood.x = replace_na(alcblood.x, 0),
         # create time (days) until event happen
         time = case_when(
           !is.na(FROM_DT) ~ as.numeric(difftime(FROM_DT, index.x, units = c("days"))), 
           is.na(FROM_DT) ~ as.numeric(difftime(clm_last, index.x,units = c("days")))),
         surgery_type = "c")
```

## Summary statistics for patients
```{r}
b_all %>% 
  select(der_sex, der_age) %>% 
  summarise(med_age = median(der_age),
            min_age = min(der_age),
            max_age = max(der_age),
            q1_age = quantile(der_age, 1/4),
            q3_age = quantile(der_age, 3/4))

b_perc_alc <- dim(b_outcom_unique)[1]/dim(b_enroll)[1]

c_all %>% 
  select(der_sex, der_age) %>% 
  summarise(med_age = median(der_age),
            min_age = min(der_age),
            max_age = max(der_age),
            q1_age = quantile(der_age, 1/4),
            q3_age = quantile(der_age, 3/4))

c_perc_alc <- dim(c_outcom_unique)[1]/dim(c_enroll)[1]
```


```{r cleaning 3}
# alc.x = current
# alc.y = previous year

all <- bind_rows(b_all, c_all)
```

## Kaplan-Meier estimate of survivor function
```{r}
# Build the standard survival object
# Kaplan Meier Survival Curve
# “+” after the time indicates censoring
km <- with(b_all, Surv(time, alc.x)) 
head(km,80)
```

```{r}
# Produce the Kaplan-Meier estimates of the probability of survival - bariatric patients
km_fit <- survfit(Surv(time, alc.x) ~ 1, data = b_all)

# Print the estimates for 1, 30, 60 and 90 days, and then every 90 days thereafter
summary(km_fit, times = c(1, 500, 1000, 1500*(1:5)))

plot(km_fit, xlab="Days", main = 'Kaplan Meyer Plot for bariatric surgery alcohol related outcome')
```

```{r}
# survival curve by surgery types
all <- all %>% 
  mutate(surgery_type = factor(surgery_type, labels = c("Bariatric", "Cholecystectomy")),
         status = ifelse(is.na(i), 1, 2)) # 1 = censored, 2= alcohol outcome

# Kaplan-Meyer
km_trt_fit <- survfit(Surv(time, alc.x) ~ surgery_type, data = all)
  
plot(km_trt_fit, 
         main = "Kaplan Meyer estimate of the survivor function for alcohol-related outcome",
         xlab = "First alcohol-related outcome",
         ylab = "Estimated survivor function",
     newdata = list(surgery_type = "Cholecystectomy"),
     col = "blue")
lines(km_trt_fit, 
         main = "Kaplan Meyer estimate of the survivor function for alcohol-related outcome",
         xlab = "First alcohol-related outcome",
         ylab = "Estimated survivor function",
     newdata = list(surgery_type = "Bariatric"),
     col = "red")
     

plot(predict(exp_plot,
             newdata = list(surgery_type = "Cholecystectomy"),
             type = "quantile",
             p = seq(.01,.99, by=.01)),
     seq(.99,.01,by = -.01), 
     col = "blue",
     type = "l",
     xlab = "Time",
     ylab = "Estimated survivor function")
lines(predict(exp_plot, 
              newdata = list(surgery_type = "Bariatric"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "red")
legend("topright", legend = c("Bariatric", "Cholecystectomy"),
       col = c("red", "blue"), lty = 1:1, cex=0.8)

# Nelson-Aalen
na_fit <- survfit(Surv(time, alc.x) ~ surgery_type, type="fh", data = all)
autoplot(na_fit, 
         main = "Nelson-Aalen estimate of the survivor function for alcohol-related outcome",
         xlab = "First alcohol-related outcome",
         ylab = "Estimated survivor function") + 
  theme_minimal() 
```

```{r}
# Bariatric
b_all_agewise <- b_all %>% 
  # Create age categorical variable
  mutate(AG = ifelse((der_age > 50), "OV60", "LT60"),
         AG = factor(AG))
         # prior = factor(alc.y))

km_AG_fit <- survfit(Surv(time, alc.x) ~ AG, data = b_all_agewise)

autoplot(km_AG_fit, main = "Kaplan Meyer Plot by Age in Bariatric invididuals")

# Cholecystectomy
c_all_agewise <- c_all %>% 
  # Create age categorical variable
  mutate(AG = ifelse((der_age > 50), "OV60", "LT60"),
         AG = factor(AG))
         #prior = factor(alc.y))

km_AG_fit <- survfit(Surv(time, alc.x) ~ AG, data = c_all_agewise)

autoplot(km_AG_fit, main = "Kaplan Meyer Plot by Age in Cholecystectomy invididuals")
```


## Common Cox proportional model

*Cumulative hazards*

```{r}
# exponential fit
all <- all %>% 
  filter(time > 0)

exp_plot <- survreg(Surv(time, alc.x) ~ surgery_type, data = all, dist="exponential")
plot(predict(exp_plot,
             newdata = list(surgery_type = "Cholecystectomy"),
             type = "quantile",
             p = seq(.01,.99, by=.01)),
     seq(.99,.01,by = -.01), 
     col = "blue",
     type = "l",
     xlab = "Time",
     ylab = "Estimated survivor function")
lines(predict(exp_plot, 
              newdata = list(surgery_type = "Bariatric"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "red")
legend("topright", legend = c("Bariatric", "Cholecystectomy"),
       col = c("red", "blue"), lty = 1:1, cex=0.8)

# weibull fit
weib_plot <- survreg(Surv(time, alc.x) ~ surgery_type, data = all, dist="weibull")
plot(predict(weib_plot,
             newdata = list(surgery_type = "Cholecystectomy"),
             type = "quantile",
             p = seq(.01,.99, by=.01)),
     seq(.99,.01,by = -.01), 
     col = "blue",
     type = "l",
     xlab = "Time",
     ylab = "Estimated survivor function")
lines(predict(weib_plot, 
              newdata = list(surgery_type = "Bariatric"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "red")
legend("topright", legend = c("Bariatric", "Cholecystectomy"),
       col = c("red", "blue"), lty = 1:1, cex=0.8)
```

```{r}
# create a survival object
with(b_all, Surv(depression, alc.x))
my.fit <- survfit(Surv(depression, alc.x) ~ 1, data = b_all)
names(my.fit)

# hazard = -log(S)
H.hat <- -log(my.fit$surv)
H.hat <- c(H.hat, tail(H.hat, 1)) 
H.hat[!is.infinite(H.hat)]

# A summary plot or table created using H.hat with my.fit$time
h.sort.of <- my.fit$n.event / my.fit$n.risk
H.tilde <- cumsum(h.sort.of)
H.tilde <- c(H.tilde, tail(H.tilde, 1))
 
# plotting
plot(c(my.fit$time, 250), H.hat, xlab="time", ylab="cumulative hazard",
main="comparing cumulative hazards")
# ylim=range(c(H.hat, H.tilde)), type="s")
points(c(my.fit$time, 250), H.tilde, lty=2, type="s")
legend("topleft", legend=c("H.hat","H.tilde"), lty=1:2)
```


```{r}
# patient that have alcohol outcome
# count combination of patient_id and FROM_DT, no recurrences?
alc.all <- all %>%
  # select(pat_id, FROM_DT, index.y, alc.x, surgery_type, time) %>% 
  filter(!is.na(index.y)) %>% 
  group_by(pat_id) %>% 
  count(FROM_DT) %>% 
  filter(n > 1)
```

https://shariq-mohammed.github.io/files/cbsa2019/1-intro-to-survival.html
```{r}
# cox fit 
fit <- survfit(Surv(time, alc.x) ~ surgery_type, data = all)
ggsurvplot(fit, data = all, risk.table = TRUE, size = 1,
           legend = "top",
           # pval = TRUE,
           linetype = "strata",
           fun = "pct",
           legend.title = "Surgery type",
           legend.labs = c("Bariatric Surgery",
                           "Cholecystectomy"))

# cum hazard
ggsurvplot(fit, data = all, fun = "cumhaz", legend.title = "Surgery type")

# table summary
coxph(Surv(time, alc.x) ~ surgery_type, data = all) %>% 
  gtsummary::tbl_regression(exp = TRUE)

# diagnostic plot 
# PH assumption
# ggcoxzph(cox.zph(cox.fit))
# residuals
ggcoxdiagnostics(cox.fit, type = "deviance",
                 linear.predictions = FALSE, 
                 ggtheme = theme_bw())
```

```{r}
# patient that have alcohol outcome
ggplot(all, aes(x = der_age, y = time)) +
    geom_point(aes(color = factor(surgery_type), alpha = 0.2))+
    ggtitle("Censored obs. in red") +
    theme_bw()
```

```{r}
# weibull distribution
weibull.all <- all %>% 
  filter(time != 0)

(fit <- survreg(Surv(time, alc.x) ~ der_age, data = weibull.all))
```

```{r}
# longest follow-up period
b_all_temp <- b_all %>% 
  mutate(period = clm_last - clm_frst)
max(b_all_temp$period)
c_all_temp <- c_all %>% 
  mutate(period = clm_last - clm_frst)
max(c_all_temp$period)
```

