---
title: "Bariatric_surgery"
author: "Yihan Shi"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load library, cache = TRUE}
library(survival)
library(ranger)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggfortify)
library(devtools)
library(OIsurv)
library(lubridate)
```

```{r load data, cache = TRUE}
data <- load("recur.RData")
data
```

## Data dictionary
In *x_outcome* dataset, 
Claim-level information - only claims which correspond to an alcohol-related outcome
alc: alcohol outcomes
FROM_DT: date of the claim (alcohol outcome), outcome for each individual patient (PAT_ID) can be tracked through date

*If days are consecutive, count as 1 occurrence*

*b_enroll* dataset contains patient level characteristics
# index = fist relevavnt surgery date
```{r bariatric surgery}
b_outcom # 1361 unique rows
b_expsum
b_enroll # 48997 rows
```

*c_expsum* dataset contains patient level characteristics for the year prior to the indexed surgery date
sums of how many times patients had a claim for various conditions (may adjust for baseline)

```{r control surgery }
c_outcom
c_expsum
c_enroll
```

## Cleaning
```{r cleaning b}
# Make a data set with unique combinations of patient ID & FROM_DT for outcom
b_outcom_unique <-  b_outcom %>% 
  filter(!duplicated(PAT_ID,FROM_DT)) 

# check NA values in a column
sum(is.na(b_outcom_unique$FROM_DT))

# Join enroll, outcom, and expsum
b_outcome_enroll <- b_enroll %>% 
  left_join(b_outcom_unique, c("pat_id" = "PAT_ID"))

# check NA values in a column
dim(b_outcome_enroll)[1] - sum(is.na(b_outcome_enroll$FROM_DT))

b_all <- b_outcome_enroll %>% 
  left_join(b_expsum, c("pat_id" = "PAT_ID")) %>% 
  mutate(der_age = 2021 - der_yob,
         alc.x = replace_na(alc.x, 0),
         alcabuse.x = replace_na(alcabuse.x, 0),
         alcdep.x = replace_na(alcdep.x, 0),
         alcgastr.x = replace_na(alcgastr.x, 0),
         ald.x = replace_na(ald.x, 0),
         alcpois.x = replace_na(alcpois.x, 0),
         alcsuic.x = replace_na(alcsuic.x, 0),
         alcblood.x = replace_na(alcblood.x, 0),
         # create time (days) until event happen
         time = case_when(
           !is.na(FROM_DT) ~ difftime(FROM_DT, clm_frst, units = c("days")),
           is.na(FROM_DT) ~ difftime(clm_last, clm_frst,units = c("days"))
         ),
         surgery_type = "b")

# sum(b_enroll$pat_id %in% b_all$pat_id)

# dim(b_outcom_unique)
# dim(b_outcom)
```


```{r cleaning c}
c_outcom_unique <-  c_outcom %>% 
  filter(!duplicated(PAT_ID,FROM_DT))

# Join enroll, outcom, and expsum
c_outcome_enroll <- c_enroll %>% 
  left_join(c_outcom_unique, c("pat_id" = "PAT_ID"))

c_all <- c_outcome_enroll %>% 
  left_join(c_expsum, c("pat_id" = "PAT_ID")) %>% 
  mutate(der_age = 2021 - der_yob,
         alc.x = replace_na(alc.x, 0),
         alcabuse.x = replace_na(alcabuse.x, 0),
         alcdep.x = replace_na(alcdep.x, 0),
         alcgastr.x = replace_na(alcgastr.x, 0),
         ald.x = replace_na(ald.x, 0),
         alcpois.x = replace_na(alcpois.x, 0),
         alcsuic.x = replace_na(alcsuic.x, 0),
         alcblood.x = replace_na(alcblood.x, 0),
         # create time (days) until event happen
         time = case_when(
           !is.na(FROM_DT) ~ difftime(FROM_DT, clm_frst, units = c("days")),
           is.na(FROM_DT) ~ difftime(clm_last, clm_frst,units = c("days"))
         ),
         surgery_type = "c")

sum(c_enroll$pat_id %in% c_all$pat_id)
```

## Summary statistics for patients
```{r}
b_all %>% 
  select(der_sex, der_age) %>% 
  summarise(med_age = median(der_age),
            min_age = min(der_age),
            max_age = max(der_age),
            q1_age = quantile(der_age, 1/4),
            q3_age = quantile(der_age, 3/4))

b_perc_alc <- dim(b_outcom_unique)[1]/dim(b_enroll)[1]

c_all %>% 
  select(der_sex, der_age) %>% 
  summarise(med_age = median(der_age),
            min_age = min(der_age),
            max_age = max(der_age),
            q1_age = quantile(der_age, 1/4),
            q3_age = quantile(der_age, 3/4))

c_perc_alc <- dim(c_outcom_unique)[1]/dim(c_enroll)[1]
```


```{r cleaning 3}
# alc.x = current
# alc.y = previous year

all <- bind_rows(b_all, c_all)
```

## Kaplan-Meier estimate of survivor function
```{r}
# Build the standard survival object
# Kaplan Meier Survival Curve
# “+” after the time indicates censoring
km <- with(b_all, Surv(time, alc.x))
head(km,80)
```

```{r}
# Produce the Kaplan-Meier estimates of the probability of survival 
km_fit <- survfit(Surv(time, alc.x) ~ 1, data = b_all)

# Print the estimates for 1, 30, 60 and 90 days, and then every 90 days thereafter
summary(km_fit, times = c(1,500,1000,1500*(1:5)))

plot(km_fit, xlab="Days", main = 'Kaplan Meyer Plot for bariatric surgery related outcome')
```

```{r}
# survival curve by surgery type
all <- all %>% 
  mutate(surgery_type = factor(surgery_type, labels = c("Bariatric", "Cholecystectomy")))

km_trt_fit <- survfit(Surv(time, alc.x) ~ surgery_type, data = all)
autoplot(km_trt_fit, main = "Kaplan Meyer Plot by Surgery type")
```

```{r}
# Survival by age
# same plot for b & c?

# Bariatric
b_all_agewise <- b_all %>% 
  # Create age categorical variable
  mutate(AG = ifelse((der_age > 50), "OV60", "LT60"),
         AG = factor(AG))
         #prior = factor(alc.y))

km_AG_fit <- survfit(Surv(time, alc.x) ~ AG, data = b_all_agewise)

autoplot(km_AG_fit, main = "Kaplan Meyer Plot by Age in Bariatric invididuals")

# Cholecystectomy
c_all_agewise <- c_all %>% 
  # Create age categorical variable
  mutate(AG = ifelse((der_age > 50), "OV60", "LT60"),
         AG = factor(AG))
         #prior = factor(alc.y))

km_AG_fit <- survfit(Surv(time, alc.x) ~ AG, data = b_all_agewise)

autoplot(km_AG_fit, main = "Kaplan Meyer Plot by Age in Cholecystectomy invididuals")
```


## Common Cox proportional model

*Cumulative hazards*

```{r}
# create a survival object
with(b_all, Surv(depression, alc.x))
my.fit <- survfit(Surv(depression, alc.x) ~ 1, data = b_all)
names(my.fit)

# hazard = -log(S)
H.hat <- -log(my.fit$surv)
H.hat <- c(H.hat, tail(H.hat, 1)) 
H.hat[!is.infinite(H.hat)]

# A summary plot or table created using H.hat with my.fit$time
h.sort.of <- my.fit$n.event / my.fit$n.risk
H.tilde <- cumsum(h.sort.of)
H.tilde <- c(H.tilde, tail(H.tilde, 1))
 
# plotting
plot(c(my.fit$time, 250), H.hat, xlab="time", ylab="cumulative hazard",
main="comparing cumulative hazards")
# ylim=range(c(H.hat, H.tilde)), type="s")
points(c(my.fit$time, 250), H.tilde, lty=2, type="s")
legend("topleft", legend=c("H.hat","H.tilde"), lty=1:2)
```
