---
title: "clean_analysis"
author: "Yihan Shi"
date: "11/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load library, cache = TRUE}
library(survival)
library(survminer)
library(ranger)
library(gt)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggfortify)
library(devtools)
# library(OIsurv)
library(lubridate)
```
 
Pat_id: ID
Enr_frst: date of first enrollment in coverage
Enr_last: date of last enrollment in coverage
Der_sex: sex
Der_yob: year of birth
Pat_region: census region of patient
Gbp: indicator for gastric bypass
Slg: indicator for sleeve gastrectomy
Agb: indicator for adjustable gastric band
Dds: indicator for duodenal switch
Depression: indicator for presence of depression prior to surgery (one year window)
Anxiety: indicator for presence of anxiety prior to surgery (one year window)
Cci: Charlson comorbidity index (one year window prior to surgery)
FROM_DT: date of alcohol outcome – NA means that this patient did not experience an outcome and their event type should be a censoring at their enr_last
Alcabuse: indicator for alcohol outcome (in this case, specifically a claim related to acute alcohol abuse)

```{r load data, cache = TRUE}
data <- load("/Users/yihanshi/Desktop/Duke_2021-22/Jiang research/Survival_analysis_practice/data/recur_211127.RData")
data2 <- load("/Users/yihanshi/Desktop/Duke_2021-22/Jiang research/Survival_analysis_practice/data/recur.RData")
data
data2
b_outcom # con_num not missing = hospital claim
```


```{r merge}
bariatric <- bariatric %>% 
  inner_join(b_enroll, c("pat_id" = "pat_id")) %>% 
  select("pat_id", "enr_frst.x", "enr_last.x", "der_sex.x", "der_yob.x", "pat_region.x",
         "gbp", "slg", "agb", "dds", "depression", "anxiety", "cci", "FROM_DT", "alcabuse",
         "index", "clm_last") %>% 
  rename(enr_frst = enr_frst.x,
         enr_last = enr_last.x,
         der_sex = der_sex.x,
         der_yob = der_yob.x,
         pat_region = pat_region.x)

chole <- chole %>% 
  inner_join(c_enroll, c("pat_id" = "pat_id")) %>% 
  select("pat_id", "enr_frst.x", "enr_last.x", "der_sex.x", "der_yob.x", "pat_region.x",
         "depression", "anxiety", "cci", "FROM_DT", "alcabuse",
         "index") %>% 
  rename(enr_frst = enr_frst.x,
         enr_last = enr_last.x,
         der_sex = der_sex.x,
         der_yob = der_yob.x,
         pat_region = pat_region.x)
```


```{r summary}
bariatric <- bariatric %>%
  mutate(der_age = 2021 - der_yob,
         time = case_when(
           !is.na(FROM_DT) ~ as.numeric(difftime(FROM_DT, index, units = c("days"))), 
           is.na(FROM_DT) ~ as.numeric(difftime(enr_last, index, units = c("days")))),
         censored = ifelse(is.na(FROM_DT), 1, 0)) %>%  # 1 = censored/no outcome, 0 = alcohol outcome
  filter(time > 0)

# age distribution
bariatric %>% 
  select(der_sex, der_age) %>% 
  summarise(med_age = median(der_age),
            min_age = min(der_age),
            max_age = max(der_age),
            q1_age = quantile(der_age, 1/4),
            q3_age = quantile(der_age, 3/4))

# percentage alc 
alc_count <- bariatric %>% 
  count(is.na(FROM_DT)) %>% 
  mutate(freq = n/sum(n)) # 4.8% experience alcohol outcome

# by surgery types
b_longer <- bariatric %>% 
  pivot_longer(
    cols = c("gbp", "slg", "agb", "dds"), # there is no dds patients
    names_to = "type") %>% 
  filter(value == 1)
```

```{r check how many have experienced alc}
ds <- b_longer %>% 
  group_by(censored) %>% 
  count()
```


## KM

```{r}
# Produce the Kaplan-Meier estimates of the probability of survival - bariatric patients
km_fit <- survfit(Surv(time, censored) ~ type, data = b_longer)

# Print the estimates for 1, 30, 60 and 90 days, and then every 90 days thereafter
summary(km_fit, times = c(1, 500, 1000, 1500*(1:5)))

plot(km_fit, 
     main = 'Kaplan Meyer Plot for bariatric surgery alcohol related outcome',
     xlab = "First alcohol-related outcome",
     ylab = "Survival Probabilities",
     col = c("red", "dark green", "blue"),
     lty = c("solid", "dashed", "dotted"))
legend("topright", 
       c("agb", "gbp", "slg"), 
       col = c("red", "dark green", "blue"),
       lty = c("solid", "dashed", "dotted"))
```

## NA

```{r}
na_fit <- survfit(Surv(time, censored) ~ type, type = "fh", data = b_longer)
plot(na_fit, 
     main = 'Nelson-Aalen Plot for bariatric surgery alcohol related outcome',
     xlab = "First alcohol-related outcome",
     ylab = "Survival Probabilities",
     col = c("red", "dark green", "blue"),
     lty = c("solid", "dashed", "dotted"))
legend("topright", 
       c("agb", "gbp", "slg"), 
       col = c("red", "dark green", "blue"),
       lty = c("solid", "dashed", "dotted"))
# autoplot(na_fit, 
#          main = "Nelson-Aalen estimate of the survivor function for alcohol-related outcome",
#          xlab = "First alcohol-related outcome",
#          ylab = "Estimated survivor function") + 
#   theme_minimal() 
```

## Common Cox proportional model

*Cumulative hazards*

```{r}
# exponential fit
exp_plot <- survreg(Surv(time, censored) ~ type, data = b_longer, dist = "exponential")
plot(predict(exp_plot,
             newdata = list(type = "slg"),
             type = "quantile",
             p = seq(.01,.99, by=.01)),
     seq(.99,.01,by = -.01), 
     col = "blue",
     type = "l",
     main = "Cox proportional model based on exponential distribution",
     xlab = "Time",
     ylab = "Estimated survivor function",
     lty = "dotted")
lines(predict(exp_plot, 
              newdata = list(type = "gbp"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "dark green",
      lty = "dashed")
lines(predict(exp_plot, 
              newdata = list(type = "agb"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "red",
      lty = "solid")
legend("topright", 
       c("agb", "gbp", "slg"), 
       col = c("red", "dark green", "blue"),
       lty = c("solid", "dashed", "dotted"))
```

```{r}
# weibull fit
weib_plot <- survreg(Surv(time, censored) ~ type, data = b_longer, dist = "weibull")
plot(predict(weib_plot,
             newdata = list(type = "slg"),
             type = "quantile",
             p = seq(.01,.99, by=.01)),
     seq(.99,.01,by = -.01), 
     col = "blue",
     type = "l",
     main = "Cox proportional model based on exponential distribution",
     xlab = "Time",
     ylab = "Estimated survivor function",
     lty = "dotted")
lines(predict(weib_plot, 
              newdata = list(type = "gbp"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "dark green",
      lty = "dashed")
lines(predict(weib_plot, 
              newdata = list(type = "agb"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "red",
      lty = "solid")
legend("topright", 
       c("agb", "gbp", "slg"), 
       col = c("red", "dark green", "blue"),
       lty = c("solid", "dashed", "dotted"))
```

## Cox fit
```{r}
# cox fit 
ggsurvplot(km_fit, data = b_longer, risk.table = TRUE, size = 1,
           legend = "top",
           linetype = "strata",
           fun = "pct",
           legend.title = "Surgery type",
           legend.labs = c("agb", "gbp", "slg"),
           risk.table.height = 0.32)

# cum hazard
ggsurvplot(km_fit, data = b_longer, fun = "cumhaz", legend.title = "Surgery type")

# table summary
coxph(Surv(time, censored) ~ type, data = b_longer) %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

## Recurrent event models

```{r test}
# tstart/tstop for an individual
test <-  b_longer %>% 
  filter(pat_id == "0268AAAAAABHWZFP") %>%
  mutate(diff = time - lag(time)) %>% 
  mutate(diff = ifelse(censored == 0 & is.na(diff), 0, diff)) 

         
test <- test %>% 
  mutate(n = 1:n()) %>% 
  mutate(tstop = ifelse(n == 1, test[n+1,]$diff - 1, 0),
         tstart = 0) %>% 
  select("pat_id", "time", "diff", "n", "tstart", "tstop", "enr_last", "index")

for(i in 2:dim(test)[1]){
    test$tstart[i] <- test[i-1,]$tstop + 1
    test$tstop[i] <- test[i,]$tstart + test[i+1,]$diff - 1
}

# test[dim(test)[1], "tstart"] <- test[dim(test)[1]-1,]$tstop + 1
test$tstop[dim(test)[1]] <- as.numeric(difftime(test$enr_last[dim(test)[1]], 
                                                test$index[dim(test)[1]], 
                                                units = c("days")))

test
```

```{r set tstart and tstop - 3 patients, sample dataset}
# tstart/tstop for an individual
test <-  b_longer %>% 
  filter(pat_id %in% c("0268AAAAAABHWZFP", 
                       "0309AAAAAAAAFGLS", 
                       "0309AAAAAAAASPAO")) %>%
  group_by(pat_id) %>% 
  mutate(diff = time - lag(time)) %>% 
  mutate(diff = ifelse(censored == 0 & is.na(diff), 0, diff)) 

# create a list of unique patient ids
a_test <- unique(test$pat_id)

# create empty dataframe
output <- data.frame(matrix(ncol = 8, nrow = 0))
x <- c("pat_id", "time", "diff", "n", "tstart", "tstop", "enr_last", "index")
colnames(output) <- x

for(i in 1:length(unique(a_test))){
  # create a separate df for 1 patient
  df <- test %>% 
    filter(pat_id == unique(a_test)[i])
  
  df <- df %>% 
  mutate(n = 1:n()) %>% 
  mutate(tstop = ifelse(n == 1, df[n+1,]$diff - 1, 0),
         tstart = 0) %>% 
  select("pat_id", "time", "diff", "n", "tstart", "tstop", "enr_last", "index")
  
  for(j in 2:dim(df)[1]){
    df$tstart[j] <- df[j-1,]$tstop + 1
    df$tstop[j] <- df[j,]$tstart + df[j+1,]$diff - 1
  }
  
  df$tstop[dim(df)[1]] <- as.numeric(difftime(df$enr_last[dim(df)[1]], 
                                                df$index[dim(df)[1]], 
                                                units = c("days")))
  output <- rbind(output, df)
}

output 
example <- output %>% 
  select("pat_id", "n", "tstart", "tstop") %>% 
  rename(event = n)
```

```{r set tstart and tstop - recurrent patients}
# patients who experienced recurrent alc outcome
b_recurrent <-  b_longer %>% 
  group_by(pat_id) %>% 
  filter(censored == 0 & n() > 1) %>% 
  mutate(diff = time - lag(time)) %>% 
  mutate(diff = ifelse(censored == 0 & is.na(diff), 0, diff)) 

# create a list of unique patient ids
patients <- unique(b_recurrent$pat_id)

# create empty dataframe
output_rec <- data.frame(matrix(ncol = 8, nrow = 0))
x <- c("pat_id", "time", "diff", "n", "tstart", "tstop", "enr_last", "index")
colnames(output_rec) <- x

for(i in 1:length(unique(patients))){
  # create a separate df for 1 patient
  df <- b_recurrent %>% 
    filter(pat_id == unique(patients)[i])
  
  df <- df %>% 
  mutate(n = 1:n()) %>% 
  mutate(tstop = ifelse(n == 1, df[n+1,]$diff - 1, 0),
         tstart = 0) %>% 
  select("pat_id", "time", "diff", "n", "tstart", "tstop", "enr_last", "index")
  
  for(j in 2:dim(df)[1]){
    df$tstart[j] <- df[j-1,]$tstop + 1
    df$tstop[j] <- df[j,]$tstart + df[j+1,]$diff - 1
  }
  
  df$tstop[dim(df)[1]] <- as.numeric(difftime(df$enr_last[dim(df)[1]], 
                                                df$index[dim(df)[1]], 
                                                units = c("days")))
  
  output_rec <- rbind(output_rec, df)
}

output_rec <- output_rec %>% 
  select("pat_id", "n", "tstart", "tstop")
```

```{r set tstart and tstop - non-recurrent patients & non-alc}
b_nonrecurrent <-  b_longer %>% 
  group_by(pat_id) %>% 
  filter(censored == 0 & n() == 1) %>% 
  mutate(tstart = 0,
         tstop = as.numeric(difftime(FROM_DT, index, units = c("days"))),
         n = 1) %>% 
  select("pat_id", "n", "tstart", "tstop")

b_nonalc <- b_longer %>% 
  group_by(pat_id) %>% 
  filter(censored == 1) %>% 
  mutate(tstart = 0,
         tstop = as.numeric(difftime(enr_last, index, units = c("days"))),
         n = 1) %>% 
  select("pat_id", "n", "tstart", "tstop")

# 2 patients have stop time of -1
output_all <- rbind(output_rec, b_nonrecurrent, b_nonalc) %>% 
  filter(tstop > 0 & (tstart < tstop))

# add surgery type and censored data
recur_data <- output_all %>% 
  inner_join(b_longer, c("pat_id" = "pat_id")) %>% 
  distinct(pat_id, tstart, tstop, censored, type, .keep_all = TRUE) %>% 
  group_by(pat_id)
```



1. Andersen-Gill (AG) model
```{r AG}
model_AG <- coxph(Surv(tstart, tstop, censored) ~ type + cluster(pat_id), 
                 method = "breslow", 
                 cluster = pat_id,
                 robust = TRUE, data = recur_data)
summary(model_AG) 

# plot AG - error
plot(survfit(model_AG))

# table summary
model_AG %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

```{r PWP-TT model}
# PWP-TT model
model_pwptt <- coxph(Surv(tstart, tstop, censored) ~ type + cluster(pat_id) + strata(n),
                     method = "breslow",
                     data = recur_data)
summary(model_pwptt)

model_pwptt %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r PWP-GT models}
# PWP-GT model
pwpgt_recur_data <- recur_data %>% 
  select("pat_id", "tstart", "tstop", "censored", "n")

# error: invalid 'times' argument
# model_pwpgt <- coxph(Surv(rep(0,dim(data)[2]), 
#                           tstop - start, 
#                           censored) ~ type * strata(n) + cluster(pat_id), 
#                      method = "breslow",  data = pwpgt_recur_data)
summary(model_pwpgt)
```
```{r WLW model}

```


