---
title: "clean_analysis"
author: "Yihan Shi"
date: "11/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load library, cache = TRUE}
library(survival)
library(survminer)
library(ranger)
library(gt)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggfortify)
library(devtools)
# library(OIsurv)
library(lubridate)
```
 
Pat_id: ID
Enr_frst: date of first enrollment in coverage
Enr_last: date of last enrollment in coverage
Der_sex: sex
Der_yob: year of birth
Pat_region: census region of patient
Gbp: indicator for gastric bypass
Slg: indicator for sleeve gastrectomy
Agb: indicator for adjustable gastric band
Dds: indicator for duodenal switch
Depression: indicator for presence of depression prior to surgery (one year window)
Anxiety: indicator for presence of anxiety prior to surgery (one year window)
Cci: Charlson comorbidity index (one year window prior to surgery)
FROM_DT: date of alcohol outcome – NA means that this patient did not experience an outcome and their event type should be a censoring at their enr_last
Alcabuse: indicator for alcohol outcome (in this case, specifically a claim related to acute alcohol abuse)

```{r load data, cache = TRUE}
data <- load("recur_211127.RData")
data
```

```{r}
head(bariatric)
head(i)
```
```{r summary}
bariatric <- bariatric%>%
  mutate(der_age = 2021 - der_yob,
         time = case_when(
           !is.na(FROM_DT) ~ as.numeric(difftime(FROM_DT, enr_frst, units = c("days"))), 
           is.na(FROM_DT) ~ as.numeric(difftime(enr_last, enr_frst, units = c("days")))),
         censored = ifelse(is.na(FROM_DT), 1, 0)) # 1 = censored/no outcome, 0 = alcohol outcome

# age distribution
bariatric %>% 
  select(der_sex, der_age) %>% 
  summarise(med_age = median(der_age),
            min_age = min(der_age),
            max_age = max(der_age),
            q1_age = quantile(der_age, 1/4),
            q3_age = quantile(der_age, 3/4))

# percentage alc 
alc_count <- bariatric %>% 
  count(is.na(FROM_DT)) %>% 
  mutate(freq = n/sum(n)) # 4.8% experience alcohol outcome

# by surgery types
b_longer <- bariatric %>% 
  pivot_longer(
    cols = c("gbp", "slg", "agb", "dds"), # there is no dds patients
    names_to = "type") %>% 
  filter(value == 1)
```

## KM

```{r}
# Produce the Kaplan-Meier estimates of the probability of survival - bariatric patients
km_fit <- survfit(Surv(time, censored) ~ type, data = b_longer)

# Print the estimates for 1, 30, 60 and 90 days, and then every 90 days thereafter
summary(km_fit, times = c(1, 500, 1000, 1500*(1:5)))

plot(km_fit, 
     main = 'Kaplan Meyer Plot for bariatric surgery alcohol related outcome',
     xlab = "First alcohol-related outcome",
     ylab = "Survival Probabilities",
     col = c("blue", "dark green", "red"),
     lty = c("solid", "dashed", "dotted"))
legend("topright", 
       c("slg", "gbp", "agb"), 
       col = c("blue", "dark green", "red"),
       lty = c("solid", "dashed", "dotted"))
```

## NA

```{r}
na_fit <- survfit(Surv(time, censored) ~ type, type = "fh", data = b_longer)
autoplot(na_fit, 
         main = "Nelson-Aalen estimate of the survivor function for alcohol-related outcome",
         xlab = "First alcohol-related outcome",
         ylab = "Estimated survivor function") + 
  theme_minimal() 
```

## Common Cox proportional model

*Cumulative hazards*

```{r}
# exponential fit
exp_plot <- survreg(Surv(time, censored) ~ type, data = b_longer, dist = "exponential")
plot(predict(exp_plot,
             newdata = list(type = "slg"),
             type = "quantile",
             p = seq(.01,.99, by=.01)),
     seq(.99,.01,by = -.01), 
     col = "blue",
     type = "l",
     xlab = "Time",
     ylab = "Estimated survivor function",
     lty = "solid")
lines(predict(exp_plot, 
              newdata = list(type = "gbp"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "dark green",
      lty = "dashed")
lines(predict(exp_plot, 
              newdata = list(type = "agb"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "red",
      lty = "dotted")
legend("topright", 
       c("slg", "gbp", "agb"), 
       col = c("blue", "dark green", "red"),
       lty = c("solid", "dashed", "dotted"))
```

```{r}
# weibull fit
weib_plot <- survreg(Surv(time, censored) ~ type, data = b_longer, dist = "weibull")
plot(predict(weib_plot,
             newdata = list(type = "slg"),
             type = "quantile",
             p = seq(.01,.99, by=.01)),
     seq(.99,.01,by = -.01), 
     col = "blue",
     type = "l",
     xlab = "Time",
     ylab = "Estimated survivor function",
     lty = "solid")
lines(predict(weib_plot, 
              newdata = list(type = "gbp"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "dark green",
      lty = "dashed")
lines(predict(weib_plot, 
              newdata = list(type = "agb"),
              type="quantile", 
              p = seq(.01,.99,by = .01)), 
      seq(.99,.01,by  =-.01), 
      col = "red",
      lty = "dotted")
legend("topright", 
       c("slg", "gbp", "agb"), 
       col = c("blue", "dark green", "red"))
```
## Cox fit
```{r}
# cox fit 
ggsurvplot(km_fit, data = b_longer, risk.table = TRUE, size = 1,
           legend = "top",
           linetype = "strata",
           fun = "pct",
           legend.title = "Surgery type",
           legend.labs = c("agb", "gbp", "slg"),
           risk.table.height = 0.32)

# cum hazard
ggsurvplot(km_fit, data = b_longer, fun = "cumhaz", legend.title = "Surgery type")

# table summary
coxph(Surv(time, censored) ~ type, data = b_longer) %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

## Recurrent event models

```{r}
bariatric_alc <- b_longer %>% 
  # filter(!is.na(FROM_DT)) %>% 
  group_by(pat_id) %>% 
  arrange(pat_id, FROM_DT) %>% 
  mutate(diff = FROM_DT - lag(FROM_DT), 
         diff_days = as.numeric(diff, units = 'days'))
  
```

1. Andersen-Gill (AG) model
```{r}
model.1 <- coxph(Surv(tstart,tstop,status) ~ var1 + var2+ … + vark, method=”breslow”, robust=TRUE, data = example1)
summary(model.1)

```

